---
title: "Invoice Collection Prediction"
output: html_notebook
---
# Libraries 
```{r results='hide'}
library(readr)
library(tidyverse)
library(forcats)
library(miscset)
library(lubridate)
library(caret)
library(ranger)

```

# Load Data

```{r results='hide'}
data <- read_csv("~/Documents/workspace/invoice_to_collection_prediction/data/Keysight/sorted_data.csv")

```

# Data Processing & Exploration
```{r}
glimpse(data)
```


```{r}

data %>%
  select_if(is.numeric) %>%
  map_dfr(n_distinct) %>%
  gather() %>%
  ggplot(aes(reorder(key, -value), value)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  #scale_y_log10(breaks = c(5, 50, 250, 500, 1000, 10000, 50000)) +
  #geom_text(aes(label = value), vjust = 1.6, color = "white", size=3.5) +
  theme_minimal() +
  labs(x = "features", y = "Number of unique values") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

* We can drop sequential index that has the same unique values **X1**, **index**
* Converting categorical features **BILL_TO_COUNTRY**, **Org_Id**, **CUST_ACCOUNT_NUMBER**, **PAYMENT_CLASS**, **PAYMENT_STATUS**, **PAYMENT_TERM**, **TRANSACTION_SUB_TYPE**, and **TRANSACTION_TYPE**


```{r}
data %>%
  select(-X1, -index, -overDueDays) %>%
  mutate(BILL_TO__COUNTRY = as.factor(BILL_TO__COUNTRY),
         Org_Id = as.factor(Org_Id),
         CUST_ACCOUNT_NUMBER = as.factor(CUST_ACCOUNT_NUMBER),
         # replace the missing values (NAs) with "missing"
         CUST_ACCOUNT_NUMBER = fct_explicit_na(CUST_ACCOUNT_NUMBER, "missing"),
         PAYMENT_CLASS = as.factor(PAYMENT_CLASS), 
         PAYMENT_STATUS = as.factor(PAYMENT_STATUS),
         PAYMENT_TERM = as.factor(PAYMENT_TERM),
         TRANSACTION_SUB_TYPE = as.factor(TRANSACTION_SUB_TYPE),
         TRANSACTION_TYPE = as.factor(TRANSACTION_TYPE)
         ) -> data

```

```{r}
data %>%
  select_if(is.factor) %>%
  summary()

data %>%
  select_if(is.numeric) %>%
  summary()


data %>%
  glimpse()
```

There are some invoices with negative values. Is that normal? 

```{r}
data %>% 
  select(AMOUNT_DUE_ORIGINAL, AMOUNT_DUE_REMAINING) %>% 
  mutate(pos_due_original = ifelse(AMOUNT_DUE_ORIGINAL >= 0, 1, 0),
         pos_due_remaining = ifelse(AMOUNT_DUE_REMAINING >= 0, 1, 0)) %>%
  summarise(pos_due_original = mean(pos_due_original), pos_due_remaining = mean(pos_due_remaining))

```
Most propably the negative invoices are cancelled ones. 

```{r}

p1 <- data %>% 
  ggplot(aes(x=log(AMOUNT_DUE_ORIGINAL), y=..density..)) + 
  geom_histogram(fill='steelblue', na.rm=TRUE, bins=40) + 
  geom_density(aes(x=log(AMOUNT_DUE_ORIGINAL)), fill='orange', color='orange', alpha=0.3, na.rm=TRUE) + 
  labs(
    title = 'Distribution of  AMOUNT_DUE_ORIGINAL',
    x = 'Natural log of AMOUNT_DUE_ORIGINAL'
  )

p2 <- data %>% 
  ggplot(aes(x=log(AMOUNT_DUE_REMAINING), y=..density..)) + 
  geom_histogram(fill='steelblue', na.rm=TRUE, bins=40) + 
  geom_density(aes(x=log(AMOUNT_DUE_REMAINING)), fill='orange', color='orange', alpha=0.3, na.rm=TRUE) + 
  labs(
    title = 'Distribution of  AMOUNT_DUE_REMAINING',
    x = 'Natural log of AMOUNT_DUE_REMAINING'
  )
ggplotGrid(list(p1, p2), ncol = 2)

rm(p1)
rm(p2)

```

We will create a new coulmn to determine number of delayes of closed invoices.

```{r}

data <- data %>%
  mutate(delinquent_days = difftime(Actual_Date_Closed, Due_Date, units = "days"))

```

Adding a new feature **ageing** to calssify the delinguent invoices.

```{r}
# Geting the max delay of closed invoices.
data %>% 
  select(PAYMENT_STATUS, delinquent_days) %>% 
  filter(PAYMENT_STATUS == "CL") %>%
  summarise(max_delay = min(delinquent_days))

# Divid the delays to bukets 
data %>%
  mutate(aging = cut(as.numeric(data$delinquent_days), br = c(-Inf, -500, 0, 30, 60, 90, Inf), labels = c('overdue', 'on_time', 'i1_30', 'i31_60', 'i61_90', 'i90_plus'))) -> data

```

# Exploring Categorical Variables 
```{r}

data %>%
  mutate(BILL_TO__COUNTRY = fct_lump(BILL_TO__COUNTRY, prop = 0.03)) %>%
  mutate(CUST_ACCOUNT_NUMBER = fct_lump(CUST_ACCOUNT_NUMBER, prop = 0.03)) %>%
  mutate(Org_Id = fct_lump(Org_Id, prop = 0.03)) %>%
  mutate(PAYMENT_TERM = fct_lump(PAYMENT_TERM, prop = 0.03)) %>%
  mutate(TRANSACTION_SUB_TYPE = fct_lump(TRANSACTION_SUB_TYPE, prop = 0.03)) -> data 

ggplotGrid(ncol = 2, hight = 100,
lapply(c("BILL_TO__COUNTRY","CUST_ACCOUNT_NUMBER","Org_Id","PAYMENT_CLASS","PAYMENT_STATUS", "PAYMENT_TERM", "TRANSACTION_SUB_TYPE", 
         "aging"),
       function(col){
         ggplot(data, aes_string(col, fill = "aging")) + geom_bar() + 
           #scale_y_log10() + 
           theme(axis.text.x = element_text(angle = 15, hjust = 1))
       }))
```
```{r}
data %>%
  select(BILL_TO__COUNTRY, CUST_ACCOUNT_NUMBER, Org_Id, PAYMENT_CLASS, PAYMENT_TERM, PAYMENT_STATUS,
         TRANSACTION_TYPE, TRANSACTION_SUB_TYPE, aging) %>%
  group_by(aging) %>%
  count()
```

# Feature Engineering 

* Adding day, month of invoice creating, due, and closing date

```{r}
# data %>%
#   select(CUST_ACCOUNT_NUMBER, INVOICE_CREATED_DATE, Due_Date, Actual_Date_Closed, aging) %>%
#   filter(aging != 'overdue') %>%
#   mutate(closed_month_day = day(Actual_Date_Closed)) %>%
#   group_by(closed_month_day, CUST_ACCOUNT_NUMBER) %>%
#   tally() %>%
#   arrange(closed_month_day) %>%
#   ggplot(aes(closed_month_day, n), col = CUST_ACCOUNT_NUMBER) + geom_col()
#   

data <- data %>%
  mutate(closed_month = as.factor(month(Actual_Date_Closed)),
         closed_day = as.factor(day(Actual_Date_Closed)), 
         due_month = as.factor(month(Due_Date)), 
         due_day = as.factor(day(Due_Date)), 
         creating_month = as.factor(month(INVOICE_CREATED_DATE)),
         creating_day = as.factor(day(INVOICE_CREATED_DATE)))

```


# Prediction Models

## Prepare Training / Testing data
```{r}
data %>% 
  select(-INVOICE_NUMBER, -BILLING_PERIOD, -PAYMENT_STATUS) %>%
  arrange(INVOICE_CREATED_DATE) -> data

len = nrow(data)
tr_length = as.integer(len * 0.7)
tr_data = data[1:tr_length, ]
tst_data = data[-(1:tr_length), ]

rm(len) 
rm(tr_length)

```

```{r}
set.seed(42)
myFolds <- createFolds(tr_data$aging, k = 5)
# Create reusable trainControl object: myControl
myControl <- trainControl(
  summaryFunction = multiClassSummary,
  classProbs = TRUE, # IMPORTANT!
  verboseIter = TRUE, #don't print the progress
  savePredictions = TRUE,
  index = myFolds
)

```

### Glment Model
```{r results="hide"}
#Glmnet Model
set.seed(42)
model_glmnet <- train(
 aging ~ ., tr_data,
 method = "glmnet",
 #preProcess = c("center", "scale", "pca"),
 tuneGrid = expand.grid(
 alpha = 0:1,
 lambda = 0:10/10
 ),
 trControl = myControl
)

```
```{r}
plot(model_glmnet)
```

```{r results='hide'}
# Random Forest
set.seed(42)
model_rf <- train(
 aging ~ ., tr_data,
 method = "ranger",
 importance = 'impurity',
 #preProcess = c("center", "scale", "pca"),
 trControl = myControl
)
```

```{r}
plot(model_rf)
```

```{r results='hdie'}
# GBM
set.seed(42)
model_gbm <- train(
 aging ~ ., tr_data,
 method = "gbm",
 #preProcess = c("center", "scale", "pca"),
 trControl = myControl
)
```

```{r}
plot(model_gbm)
```

```{r results='hide'}
#SVM
set.seed(42)
model_svm <- train(
 aging ~ ., tr_data,
 method = "svmRadial",
 #preProcess = c("center", "scale", "pca"),
 trControl = myControl
)
```

```{r}
plot(model_svm)
```


# Compering different models

```{r}
# Make a list
model_list <- list(
 glmnet = model_glmnet,
 rf = model_rf,
 gbm = model_gbm,
 svm = model_svm
 )
# Collect resamples from the CV folds
resamps <- resamples(model_list)
resamps 
```

```{r}
# Summarize the results
summary(resamps)
```


```{r}
dotplot(resamps, metric = "prAUC")
```

```{r}
dotplot(resamps, metric = "logLoss")
```

```{r}
varImp(model_glmnet)
```

# Prediction Result

```{r}
p_svm <- predict(model_svm, tst_data)
p_glmnet <- predict(model_glmnet, tst_data)
p_rf <- predict(model_glmnet, tst_data)
p_gbm <- predict(model_gbm, tst_data)

(result <- data.frame(model = c("SVM", "Glmnet", "Random Forest", "GBM"), 
           accuracy = c(mean(p_svm == tst_data$aging), 
                        mean(p_glmnet == tst_data$aging), 
                        mean(p_rf == tst_data$aging), 
                        mean(p_gbm == tst_data$aging))))

```

```{r}
result %>%
  ggplot(aes(model, accuracy)) + geom_col()
```

